---
title: "00-EDA-of-BCEF"
author: "Frances Lin"
output: pdf_document
---

## Load packages

```{r}
library(tidyverse)
library(pander)
```

```{r}
library(spNNGP)
```

```{r}
library(sf) # Need to update R & RStudio 
```

## Load data 

```{r}
data(BCEF)
```


## View data

```{r}
BCEF %>% str()
```

Description: 

- response: FCH (forest canopy height)

- covariate: PTC (Landsat derived percent tree cover) 

- x: longtitude 

- y: latitude


## Check x (longtitude) & y (latitude)

```{r}
summary(BCEF$x)
```

```{r}
summary(BCEF$y)
```

Look like p.27 in the Package spNNGP paper? Sort of. 

```{r}
BCEF %>% ggplot(aes(x = x, y = y)) + geom_point() 
```

## Check response (FCH)

FCH (forest canopy height) estimates (n = 188717) come from the G-LiHT LiDAR point cloud summarized to a $13$x$13$ m grid cell size (G-LiHT: Goddard’s LiDAR, Hyperspectral & Thermal Imager 2019). 

Over each grid cell, the maximum forest canopy height (i.e., FCH) was estimated using the 100th percentile height of the point cloud. 

```{r}
summary(BCEF$FCH)
```

```{r}
length(BCEF$FCH)
```

```{r}
BCEF %>% ggplot(aes(x = FCH)) + geom_histogram()
```

```{r}
BCEF %>% ggplot(aes(x = FCH)) + geom_boxplot()
```

Combine FCH with x and y? Look even much alike now. Still way to go though... 

```{r}
BCEF %>% ggplot(aes(x = x, y = y, color = FCH)) + geom_point() 
```

Function `geom_sf()` is used to plot object `sf`. 

```{r}
# Convert to object sf 
BCEF_sf <- BCEF %>% st_as_sf(coords = c("x", "y")) 
BCEF_sf %>% class()
```

```{r}
BCEF_sf %>% str()
```

```{r}
BCEF_sf %>% st_crs()
```

Replace FCH with PTC returns similar graph. Look like "the points" are along those lines. 

```{r}
ggplot(data = BCEF) + 
  geom_sf(data = BCEF_sf) + 
  geom_point(aes(x = x, y = y, color = FCH))
```

Check project INLA. Check function geom_tile or geom_raster. 

```{r}
library(sp) # Will retire
BCEF_df <- coordinates(BCEF)
```

```{r}
BCEF_df %>% as.data.frame %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_tile(aes(fill = FCH)) + 
  scale_fill_gradient(low = "yellow", high = "blue")
```

```{r}
ggplot(BCEF, aes(x = x, y = y, fill = FCH)) +
  geom_raster()
```


## Check covariate (PTC)

```{r}
summary(BCEF$PTC)
```

```{r}
BCEF %>% ggplot(aes(x = PTC)) + geom_histogram()
```

```{r}
BCEF %>% ggplot(aes(x = PTC)) + geom_boxplot()
```

## Check holdout 

```{r}
BCEF %>% count(holdout)
```

## Subset data 

n = 105504 

```{r}
BCEF %>% filter(holdout == 0)
```

## Subset data 2

Select n = 100000 randomly from n = 188717

```{r}
length(BCEF)
```

```{r}
# Sample rows 
set.seed(1234)
n = 100000
BCEF[sample(x = nrow(BCEF), size = n), ]
```




\newpage

## Reference 

[Finley, A. O., Datta, A., & Banerjee, S. (2020). spNNGP R package for nearest neighbor Gaussian process models. arXiv preprint arXiv:2001.09111.](https://arxiv.org/pdf/2001.09111.pdf)
